%% PART I: INTRODUCTION AND BACKGROUND
\section*{Research Question \& Significance}

Drug discovery frequently begins with a property wish list: a target molecular weight, lipophilicity, polar surface area, and related descriptors that define a pharmacological profile of interest. The practical challenge is to find synthesizable molecules that match it. Synthesis-aware generative models address this directly: given a property specification, they output molecules composed of purchasable building blocks joined via validated reactions, guaranteeing synthesizability by construction \cite{bib1}. Several models operate in this paradigm, including SynNet, ChemProjector, SynFormer, and PrexSyn \cite{bib2}.

PrexSyn, the current state of the art, generates property-conditioned candidates within the Enamine REAL chemical space \cite{bib3} at high throughput ($\sim$0.26\,s per target molecule), drawing from $\sim$480,000 in-stock building blocks and 115 reaction templates. It reconstructs 94\% of molecules drawn from Enamine REAL but only 28\% from ChEMBL \cite{bib4}, a curated database of bioactive compounds with demonstrated pharmacological activity. The remaining 72\% of ChEMBL represents pharmacologically validated chemistry that lies beyond the model's reach.

This coverage gap is structural: every synthesis-aware generator is bounded by its building block library and reaction template set. Retraining on a larger chemical space shifts the boundary but does not eliminate it, and demands substantial data curation. Repeated sampling draws from the same constrained space, producing additional candidates but opening no new chemical territory. For discovery programs seeking novel scaffolds or unexplored regions of bioactive space, this constraint directly limits the generator's practical utility.

We investigate whether post-generation molecular modification expands the space of viable candidates when composed with the frontier synthesis-aware model. Fragment substitution, matched molecular pair transforms, latent-space editing, and scaffold-constrained decoration are established techniques that predate generative AI, operate outside Enamine REAL, and are available as open-source tools. We benchmark four such methods on their ability to produce synthesizable, property-conserving candidates beyond what repeated PrexSyn sampling achieves. Results are stratified by how well PrexSyn performs alone (the baseline difficulty), revealing when modification adds the most marginal value. A complementarity analysis identifies which method combinations maximize total unique coverage, informing pipeline design and tool selection for both medicinal chemistry workflows and automated agent-driven design systems. This benchmark yields two core outputs: individual method rankings by marginal value, and optimal multi-method combinations that extend the generator's chemical reach without retraining.

\section*{Context \& Datasets}

PrexSyn \cite{bib2} generates synthesis pathways from $\sim$480,000 Enamine in-stock building blocks via 115 reaction templates, conditioned on molecular property specifications. We use it as a black-box generator: given a property profile, it returns synthesizable candidates within Enamine REAL.

ChEMBL \cite{bib4} provides realistic property specifications. We extract computed property profiles from sampled bioactive molecules: ECFP4 fingerprints (bit vectors encoding circular substructure features within a fixed bond radius, widely used for measuring molecular similarity) and physicochemical descriptors (molecular weight, lipophilicity, polar surface area, hydrogen bond donors and acceptors, and rotatable bond count). These profiles serve as input to PrexSyn. The molecules themselves are not optimization targets; structural similarity to the ChEMBL source is reported only as a supplementary diagnostic (proximity to known bioactive space).

AiZynthFinder \cite{bib5}, trained on the USPTO reaction corpus \cite{bib6}, serves as our synthesizability gate. It performs retrosynthetic analysis: computational decomposition of each candidate into purchasable Enamine building blocks via known reaction steps. Enamine REAL represents a specific, commercially relevant subset of drug-like chemistry; molecules synthesizable through alternative routes, building block suppliers, or non-template reactions may not be detected.

The four modification methods draw on independent chemical knowledge: CReM \cite{bib7} uses a context-constrained fragment database mined from ChEMBL; mmpdb \cite{bib8} applies empirical matched molecular pair transformation rules; JT-VAE \cite{bib9} encodes and decodes molecules via a learned continuous representation; LibINVENT \cite{bib10} generates scaffold decorations via reinforcement learning (accessed through REINVENT4 \cite{bib11}).
