%% PART II: PROPOSED METHODS AND SETUP
\section*{Proposed Methods}

For each property specification, PrexSyn generates an initial synthesizable candidate. Four modification methods then produce structural variants from this seed, and all variants pass through identical synthesizability and property-conservation filters. A repeated-sampling baseline (additional PrexSyn draws from the same specification) tests whether simply generating more candidates from the constrained space matches or exceeds the value of applying modification methods that operate beyond it.

\textit{CReM} \cite{bib7} modifies molecules by replacing chemical fragments (side chains, ring systems, functional groups) with alternatives drawn from a fragment database mined from real ChEMBL compounds, subject to chemical validity constraints. A tunable context radius controls the edit scale: radius 1 restricts changes to small R-group substitutions (modifications to a molecule's peripheral functional groups, e.g., swapping a methyl group for a chlorine atom, a change that can alter potency or metabolic stability); radius 2 permits medium-scale modifications such as ring substitutions; radius 3 enables scaffold-level replacements including entire ring-system swaps. This hierarchy mirrors how medicinal chemists optimize leads: iteratively modifying functional groups to improve potency, selectivity, or pharmacokinetics, starting conservatively and progressing to larger structural changes as needed. We benchmark CReM at context radii 1, 2, and 3 to characterize the edit-scale/quality tradeoff within a single method.

\textit{mmpdb} \cite{bib8} applies matched molecular pair (MMP) transformation rules: single-site substitution patterns extracted from pairs of real compounds that differ by exactly one structural change (e.g., methyl $\rightarrow$ chloro, phenyl $\rightarrow$ pyridyl). Each rule carries empirical property-change annotations derived from measured data, enabling selection of transformations likely to preserve the desired property profile. These are the most conservative edits in our benchmark: one atom or functional group at a time, with known property consequences. In medicinal chemistry, such transformations are a standard tool for systematically exploring the immediate chemical neighborhood of a lead compound (a promising early-stage candidate selected for further property optimization in a drug discovery program).

\textit{JT-VAE} \cite{bib9} encodes molecules into a continuous mathematical space (a learned latent representation) by decomposing each molecule's bonded structure into a tree of chemically meaningful substructures (a junction tree, where each node represents a ring system, bond, or atom cluster). JT-VAE generates variants by perturbing the seed molecule's latent vector and decoding the result, producing edits that span medium to global structural scales. This enables scaffold hopping: replacing an entire core structure while preserving the pharmacological profile, a strategy valued in drug discovery for accessing new chemical series and navigating intellectual property constraints.

\textit{LibINVENT} \cite{bib10,bib11} preserves a molecular scaffold (the core ring system and connectivity pattern) extracted from PrexSyn's output while generating alternative R-group decorations (the functional groups attached to the scaffold) via reinforcement learning steered toward target properties. Where CReM and mmpdb edit at the fragment or single-site level, and JT-VAE operates across global structural scales, LibINVENT occupies a distinct niche: scaffold-constrained decoration. It systematically varies the peripheral chemistry that often drives selectivity and pharmacokinetic behavior, while holding the core structure fixed.

These four methods span two independent axes: the rule-based/learned axis (CReM and mmpdb apply deterministic rules from empirical databases; JT-VAE and LibINVENT use trained neural networks) and the local-to-global edit-scale axis (mmpdb $\rightarrow$ CReM $\rightarrow$ JT-VAE, with LibINVENT operating under scaffold constraints). Selection was informed by a systematic review of $\sim$50 candidate methods, prioritizing coverage across edit scales and paradigms. No prior work systematically benchmarks modification methods as post-generation complements to synthesis-aware generation. Beyond individual method rankings, a complementarity analysis identifies which method combinations maximize total unique coverage. These rankings and complementarity matrices provide the empirical basis for composing modification operators into multi-stage pipelines with synthesis filtering at every stage, a capability absent from existing tools and applicable to both medicinal chemistry workflows and automated agent-driven pipelines.
