%% Data and Preprocessing
\section*{Data and Preprocessing}

We sample 100 molecules from ChEMBL with pre-stratified difficulty. PrexSyn is first run on a larger pool to compute baseline quality scores (ECFP4 Tanimoto similarity between the best candidate and the specification fingerprint, where Tanimoto measures the fraction of shared features between two binary fingerprints on a 0-to-1 scale). Approximately 25 molecules are then drawn from each of four quality bins ($<$0.5, 0.5--0.7, 0.7--0.85, 0.85--1.0), ensuring balanced representation across difficulty levels. This pre-stratification prevents easy specifications from dominating the benchmark and guarantees statistical power in every difficulty regime.

For each specification, PrexSyn generates 256 candidates; the highest-scoring candidate ($K=1$) seeds all four modification methods, so differences in output reflect the modification method rather than seed selection. Top-$K$ seeding ($K>1$), which tests whether seed diversity compounds with method diversity, is a natural extension; we fix $K=1$ here for clean causal attribution. Each method produces up to $N$ variants; $N$ is determined by a pilot saturation study (5 specs $\times$ 4 methods $\times$ 1,000 variants) that identifies the diversity plateau beyond which additional variants yield diminishing structural novelty. Expected cap: 100--500 variants per method per specification. Variants are deduplicated by canonical SMILES (a standardized text representation of molecular structure) before synthesizability checking via AiZynthFinder at search depth 1, with depth-2 sensitivity analysis on a 10\% subsample to quantify routes missed by shallower search. Estimated total load after deduplication: $\sim$80,000 unique molecules, processable in $\sim$8 core-hours.

No models are trained in this study; all methods operate as pre-trained or pre-configured tools. CReM's fragment database derives from ChEMBL, the same source as our property specifications. Because ChEMBL molecules serve only as property-profile sources and not as reconstruction targets, this overlap reflects realistic tool usage rather than information leakage.
