%% Experimental Setup & Baselines
\section*{Experimental setup \& baselines}

\textit{Baseline.} Repeated PrexSyn sampling: $N$ additional draws from the same property specification, subject to identical evaluation. This tests whether adding a modification layer outperforms simply sampling more from the generative model. The pilot saturation study (Section ``Data and Preprocessing'') determines $N$ at or beyond each method's diversity plateau, ensuring a fair comparison: neither method nor baseline is evaluated below its saturation point.

\textit{Hit classification.} We classify a variant as a hit if it is (a) synthesizable (AiZynthFinder route found) and (b) property-conserving on both axes defined below. We treat thresholds as parameters, reporting results at Tanimoto $\geq$ 0.6, 0.7, and 0.85 (spanning lenient to stringent substructural conservation) and desirability $\geq$ 0.8, so conclusions are not artifacts of a single cutoff.

\textit{Property conservation} is assessed on two complementary axes. \textit{Substructural}: ECFP4 Tanimoto similarity between variant and specification fingerprint, measuring the fraction of shared circular substructure features. Recent benchmarks demonstrate that ECFP4-based similarity matches or exceeds pretrained molecular embeddings for property-relevant comparisons. \textit{Physicochemical}: per-descriptor deltas (MW $\pm$50\,Da, CLogP $\pm$1.0, TPSA $\pm$20\,\AA$^2$, HBD $\pm$1, HBA $\pm$2, rotatable bonds $\pm$2) mapped to 0--1 desirability scores via linear tolerance functions and aggregated by geometric mean. The geometric mean ensures that a single badly violated descriptor drives the aggregate score toward zero, penalizing unbalanced profiles more stringently than an arithmetic mean would. These are the same descriptors PrexSyn conditions on, ensuring evaluation matches the generator's abstraction level. We report 3D shape and electrostatic similarity via ESPsim \cite{bib12} as a supplementary diagnostic, not a hit-classification gate, because it introduces a dependency on the reference molecule's 3D conformer geometry (generated via the ETKDG conformer-generation algorithm) that couples evaluation to the specific ChEMBL molecule beyond its property values.

\textit{Primary metrics.} (1) \textit{Hit rate}: fraction of synthesizable variants meeting conservation thresholds; higher rates indicate more efficient use of computational budget. (2) \textit{Expansion factor}: unique hits from modification divided by unique hits from baseline, the core measure of marginal value. An expansion factor $>$1 means the method produces viable candidates inaccessible to repeated sampling alone, expanding the chemical territory available to a discovery program without retraining the generator.

\textit{Secondary metrics.} Quantitative Estimation of Drug-likeness (QED \cite{bib13}), a 0--1 composite score aggregating desirability functions over eight molecular properties (a proxy for oral bioavailability and development tractability), reported as a distribution over each method's hits to assess whether modifications preserve or degrade drug-likeness. Chemical diversity among hits ($1 - $ mean pairwise ECFP4 Tanimoto), measuring whether a method produces structurally varied candidates or clusters around a single chemotype. Wall-clock compute cost per specification. Proximity to known bioactive space (Tanimoto to ChEMBL reference molecule, reported as a diagnostic distribution only).

\textit{Complementarity.} Pairwise Jaccard overlap (shared hits divided by total unique hits across both methods) between method hit sets identifies redundant vs.\ complementary tools. We report the two- and three-method combinations that maximize total unique-hit coverage. Methods operating at different edit scales are likely to explore partially distinct chemical territory, but the degree of overlap is an empirical question: empirically measuring it determines whether multi-method pipelines are worth the added computational cost relative to using the single best-performing method alone. Low overlap indicates complementary methods; high overlap indicates redundancy. This analysis directly informs practical pipeline design: identifying the smallest set of methods that maximizes coverage guides resource allocation in both manual and automated discovery workflows.

\textit{Stratification.} We report all metrics per baseline-quality bin ($<$0.5, 0.5--0.7, 0.7--0.85, 0.85--1.0) to reveal when post-generation modification adds the most marginal value. If modification helps most when the generator performs poorly, it serves as a rescue strategy for difficult specifications; if it helps uniformly, it is a general-purpose expansion tool. This stratification directly informs how modification methods should be deployed in practice.
